---
title: "Apache Arrow"
author: "Dimitri Marinelli"
date: "25/5/2021 - #DataForGood - CorrelAid Rhein-Main"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation
In R we can install arrow from cran `install.packages("arrow")`. 
```{r}
library(arrow)
arrow_info()

# install_arrow() # to include further properties like support for S3 or compression
# install_pyarrow() # to be used with reticulate
```



## Get files from Python

We can easily load the files exchanged with both Parquet and Feather

```{r}
planets <- arrow::read_feather("data/planets.feather")
planets_arrow <- read_feather("data/planets.feather", as_data_frame = FALSE)
planets |> class()
planets_arrow |> head()
planets_arrow |> class()
```
it preserves the column types, and even the python categories are correctly 
represented as factors in R. 

```{r}
library(ggplot2)
library(lubridate)
library(tidyverse)

planets |>  ggplot()+ geom_bar(aes(x=method)) + theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

planets |> ggplot() + 
  geom_point(aes(x=year, y=mass, color=method), alpha=0.5)+
  scale_y_log10() + theme_bw()

planets |> mutate(year=lubridate::year(year))|>  group_by(year, method) |> count() |>
  ggplot(aes(x=year, y=n, fill=method)) + 
  geom_bar(position="stack", stat="identity")+ theme_bw()

```
## Operations with arrow
It is possible to perform computations in arrow directly. It can use the dplyr
grammar. However not all functions are supported. 

```{r}
p_arrow <- planets_arrow %>%   mutate(op_mass_ratio = orbital_period / mass)
p_arrow |> compute() |> collect()
```

For optimization purposes, the first line does not actually compute the operation,
`compute()` does. `collect()` instead transforms the arrow back to tibble to be used
for further visualization. 


