---
title: "From R to Py"
author: "Dimitri Marinelli"
date: "25/5/2021 - #DataForGood - CorrelAid Rhein-Main"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

To use Python in R we can use the library "reticulate".

> The **reticulate** package provides a comprehensive set of tools for
interoperability between Python and R. The package includes facilities
for:
>
>
>  - Calling Python from R in a variety of ways including R Markdown,
>    sourcing Python scripts, importing Python modules, and using Python
>    interactively within an R session.
>
>  - Translation between R and Python objects (for example, between R and
>    Pandas data frames, or between R matrices and NumPy arrays).
>
>  - Flexible binding to different versions of Python including virtual
>    environments and Conda environments.
>
> Reticulate embeds a Python session within your R session, enabling
> seamless, high-performance interoperability. If you are an R developer
> that uses Python for some of your work or a member of data science team
> that uses both languages, reticulate can dramatically streamline your
> workflow\!



## Starting up


```{r}
# install.packages("reticulate")
library(reticulate)

# We can access directly to Python statements as functions that returns
# variable that are compatible with R

# [convert = TRUE] => convert Python objects to R when appropriate this is the
# default option.
sys <- import("sys", convert = TRUE)

# dollar operator replaces the dot in python:
sys$path

os <- import("os", convert=TRUE)

# the equivalent of sys.executable
sys$executable
```

The executable shows what environment reticulate is using. Checking this variable
is useful because reticulate tends to automatize many things, and one ends up 
not knowing which Python environment reticulate is actually using. 

It is possible to be (almost) Python agnostic, avoid any command line:

If you use conda you can check what enviroments are available from R, and then
"activate" the enviroment that reticulate will use

```{r}
print(conda_list(conda = "auto"))
use_condaenv(condaenv = "ds", conda = "auto", required = FALSE)
```
that are the equivalent of 
```
conda env list
conda activate ds
``` 
## Other examples 

```{r}
my_list <- list(x =1:2 , y=3:5)
my_dictionary <- dict(my_list)
my_dictionary
```

```{r}
my_dictionary_converted <- dict(my_list, convert = TRUE)
my_dictionary_converted$x
```

```{r}
py_to_r( my_dictionary)
```


## Use pure Python

It is possible to include in Rmd files chunks of python code, and it also 
supports the rendering of plots. (You need to write `plt.show() otherwise does 
not work).

```{python}
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt

# Load the example diamonds dataset
diamonds = sns.load_dataset("diamonds")

# Draw a scatter plot while assigning point colors and sizes to different
# variables in the dataset
f, ax = plt.subplots(figsize=(6.5, 6.5))
sns.despine(f, left=True, bottom=True)
clarity_ranking = ["I1", "SI2", "SI1", "VS2", "VS1", "VVS2", "VVS1", "IF"]
s = sns.scatterplot(x="carat", y="price",
                hue="clarity", size="depth",
                palette="ch:r=-.2,d=.3_r",
                hue_order=clarity_ranking,
                sizes=(1, 8), linewidth=0,
                data=diamonds, ax=ax)
                
plt.show();
```

However, even if the chunk is in Python, one can retrieve the variables in R. 

```{r}
py$diamonds |> head()
```

Or vice versa

```{R}
r_string <- "my R string"
```

```{python}
f'This is the string from R:  "{r.r_string}"'
```


## Use Python in R console or code

```
repl_python()
my_python_string = "this is now python" 
print(my_python_string.upper())
quit
```

## Install packages 

It is convenient for somebody who wants to stick to pure R (and not only) to
install packages directly from reticulate, for example:

```{r}
# if you use conda
# conda_install(
#  envname = NULL,
#  packages,
#  forge = TRUE,
#  channel = character(),
#  pip = FALSE,
#  pip_options = character(),
#  pip_ignore_installed = FALSE,
#  conda = "auto",
#  python_version = NULL,
#  ...
#)

# or 
# py_install("pandas")

```
the equivalent of `conda install` and `pip install`. 

